---
title: "distance function"
author: "Yunqi Yang"
date: "7/29/2020"
output:
  html_document: default
  pdf_document: default
---

## Comment:

1. Tried exponential decay function and gaussian decay function with 4 different sigma values. I think sigma values and which decay function to use don't have large impact on the ROC and PR curve. 

2. However, the coefficient for 'weight' variable changes a lot when vary sigma values. We might want to use a sigma value based on the shape of the decay curve, which one looks more reasonable in terms of biology. 


```{r  }

library(ggplot2)
library(dplyr)
require(PRROC)

```


```{r  }

load("/Users/nicholeyang/Desktop/gene_level_fine_mapping/data/training.RData")

```

```{r  }

## remove NAs
dat = rbind(train_pos_all, train_neg_all)
NAs = apply(dat, 1, function(x) sum(is.na(x)))
dat = dat[NAs == 0, ]


```


## data overview
```{r  }

head(dat)

```

```{r  }

# number of observations in the dataset 
dim(dat)[1]

sum(dat$y == 1)  # positive pairs
sum(dat$y == 0)  # negative pairs 

```


## proportion of features 

```{r  }
par(mfrow = c(1,2))

# proportion of features in the positive set
tot_count = apply(dat[dat$y==1, c(3:7)], 2, function(x) sum(x == 1))
barplot(tot_count/dim(dat[dat$y==1, ])[1], las = 2, main = 'feature proportion in positive set')

# proportion of features in the negative set
tot_count = apply(dat[dat$y==0, c(3:7)], 2, function(x) sum(x == 1))
barplot(tot_count/dim(dat[dat$y==0, ])[1], ylim = c(0, 0.01), las = 2, main = 'feature proportion in negative set')

```


## play around tss feature 
```{r  }

summary(dat$tss_dist_to_snp)
summary(dat[dat$y==1, ]$tss_dist_to_snp)  # dist_to_tss in positive set 
summary(dat[dat$y==0, ]$tss_dist_to_snp)  # dist_to_tss in negative set 

```


### two decay functions: exponential decay and gaussian decay. 

the decay functions look like under the distance range we have. 

```{r  }

gaussian_decay = function(sigma, d){
  k = exp(-d^2/sigma^2/2)
  return(k)
}
  
exp_decay = function(sigma, d){
  k = exp(-d/sigma)
  return(k)
}
  
```


```{r  }

dist = seq(0, 2e7, by = 1e3)
sigmas = c(1e4, 1e5, 1e6, 1e7)

par(mfrow = c(2,2))
plot(gaussian_decay(sigmas[1], dist), type = 'l', ylab = 'weight', xlab = 'distance')
plot(gaussian_decay(sigmas[2], dist), type = 'l', ylab = 'weight', xlab = 'distance')
plot(gaussian_decay(sigmas[3], dist), type = 'l', ylab = 'weight', xlab = 'distance')
plot(gaussian_decay(sigmas[4], dist), type = 'l', ylab = 'weight', xlab = 'distance')


par(mfrow = c(2,2))
plot(exp_decay(sigmas[1], dist), type = 'l', ylab = 'weight', xlab = 'distance')
plot(exp_decay(sigmas[2], dist), type = 'l', ylab = 'weight', xlab = 'distance')
plot(exp_decay(sigmas[3], dist), type = 'l', ylab = 'weight', xlab = 'distance')
plot(exp_decay(sigmas[4], dist), type = 'l', ylab = 'weight', xlab = 'distance')

```




## exponential decay: exp(-d/sigma) 

```{r  }

weights = matrix(NA, ncol = length(sigmas), nrow = dim(dat)[1])

for (i in 1:length(sigmas)){
  weights[, i] = exp(-dat$tss_dist_to_snp/sigmas[i])
}

weights = data.frame(weights)
colnames(weights) = c('sig1', 'sig2', 'sig3', 'sig4')


dat2 = cbind(dat, weights)

```




```{r  }


ggplot(dat2, aes(x = sig1)) +
  geom_histogram(aes(color = as.factor(y), fill = as.factor(y)), 
                position = "identity", bins = 50, alpha = 0.4) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))
  
ggplot(dat2, aes(x = sig2)) +
  geom_histogram(aes(color = as.factor(y), fill = as.factor(y)), 
                position = "identity", bins = 50, alpha = 0.4) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))
  
ggplot(dat2, aes(x = sig3)) +
  geom_histogram(aes(color = as.factor(y), fill = as.factor(y)), 
                position = "identity", bins = 50, alpha = 0.4) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))


ggplot(dat2, aes(x = sig4)) +
  geom_histogram(aes(color = as.factor(y), fill = as.factor(y)), 
                position = "identity", bins = 50, alpha = 0.4) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))
  
```



### add weight to all feature logsitic regression and plot ROC/PR-AUC 

```{r  }

# split data in the same way
set.seed(1)
n = dim(dat2)[1]
indx = sample(1:n, round(2*n/3), replace = FALSE)
train = dat2[indx, ]
test = dat2[-indx, ]


```

```{r  }

# fit model and evaluate performance
fit.sig1 = glm(y ~ UTR5 + UTR3 + intron + upstream + exon + sig1, data = train, family = "binomial")
fit.sig2 = glm(y ~ UTR5 + UTR3 + intron + upstream + exon + sig2, data = train, family = "binomial")
fit.sig3 = glm(y ~ UTR5 + UTR3 + intron + upstream + exon + sig3, data = train, family = "binomial")
fit.sig4 = glm(y ~ UTR5 + UTR3 + intron + upstream + exon + sig4, data = train, family = "binomial")


summary(fit.sig1)
summary(fit.sig2)
summary(fit.sig3)
summary(fit.sig4)

```

```{r  }

## sigma1
pred.probs=predict(fit.sig1,test,type="response")
glm.pred = rep(0, length(pred.probs))
glm.pred[pred.probs>0.5]= 1
table(glm.pred, test$y)


fg <- pred.probs[test$y == 1]
bg <- pred.probs[test$y== 0]

# ROC Curve    
roc <- roc.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(roc)

# PR Curve
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(pr)

```


```{r  }

## sigma2
pred.probs=predict(fit.sig2,test,type="response")
glm.pred = rep(0, length(pred.probs))
glm.pred[pred.probs>0.5]= 1
table(glm.pred, test$y)


fg <- pred.probs[test$y == 1]
bg <- pred.probs[test$y== 0]

# ROC Curve    
roc <- roc.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(roc)

# PR Curve
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(pr)

```


```{r  }

## sigma3
pred.probs=predict(fit.sig3,test,type="response")
glm.pred = rep(0, length(pred.probs))
glm.pred[pred.probs>0.5]= 1
table(glm.pred, test$y)


fg <- pred.probs[test$y == 1]
bg <- pred.probs[test$y== 0]

# ROC Curve    
roc <- roc.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(roc)

# PR Curve
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(pr)

```


```{r  }

## sigma4
pred.probs=predict(fit.sig4,test,type="response")
glm.pred = rep(0, length(pred.probs))
glm.pred[pred.probs>0.5]= 1
table(glm.pred, test$y)


fg <- pred.probs[test$y == 1]
bg <- pred.probs[test$y== 0]

# ROC Curve    
roc <- roc.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(roc)

# PR Curve
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(pr)

```

## change to gaussian decay function: $exp(-d^2/2l^2)$

```{r  }

weights = matrix(NA, ncol = length(sigmas), nrow = dim(dat)[1])

for (i in 1:length(sigmas)){
  weights[, i] = exp(-dat$tss_dist_to_snp^2/sigmas[i]^2)
}

weights = data.frame(weights)
colnames(weights) = c('sig1', 'sig2', 'sig3', 'sig4')
```


```{r  }

dat2 = cbind(dat, weights)

```

```{r  }


ggplot(dat2, aes(x = sig1)) +
  geom_histogram(aes(color = as.factor(y), fill = as.factor(y)), 
                position = "identity", bins = 50, alpha = 0.4) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))
  
ggplot(dat2, aes(x = sig2)) +
  geom_histogram(aes(color = as.factor(y), fill = as.factor(y)), 
                position = "identity", bins = 50, alpha = 0.4) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))
  
ggplot(dat2, aes(x = sig3)) +
  geom_histogram(aes(color = as.factor(y), fill = as.factor(y)), 
                position = "identity", bins = 50, alpha = 0.4) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))

ggplot(dat2, aes(x = sig4)) +
  geom_histogram(aes(color = as.factor(y), fill = as.factor(y)), 
                position = "identity", bins = 50, alpha = 0.4) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))
  


```


### add weight to all feature logsitic regression and plot ROC/PR-AUC 

```{r  }

# split data in the same way
set.seed(1)
n = dim(dat2)[1]
indx = sample(1:n, round(2*n/3), replace = FALSE)
train = dat2[indx, ]
test = dat2[-indx, ]


```

```{r  }

# fit model and evaluate performance
fit.sig1 = glm(y ~ UTR5 + UTR3 + intron + upstream + exon + sig1, data = train, family = "binomial")
fit.sig2 = glm(y ~ UTR5 + UTR3 + intron + upstream + exon + sig2, data = train, family = "binomial")
fit.sig3 = glm(y ~ UTR5 + UTR3 + intron + upstream + exon + sig3, data = train, family = "binomial")
fit.sig4 = glm(y ~ UTR5 + UTR3 + intron + upstream + exon + sig4, data = train, family = "binomial")


summary(fit.sig1)
summary(fit.sig2)
summary(fit.sig3)
summary(fit.sig4)

```

```{r  }

## sigma1
pred.probs=predict(fit.sig1,test,type="response")
glm.pred = rep(0, length(pred.probs))
glm.pred[pred.probs>0.5]= 1
table(glm.pred, test$y)


fg <- pred.probs[test$y == 1]
bg <- pred.probs[test$y== 0]

# ROC Curve    
roc <- roc.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(roc)

# PR Curve
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(pr)

```


```{r  }

## sigma2
pred.probs=predict(fit.sig2,test,type="response")
glm.pred = rep(0, length(pred.probs))
glm.pred[pred.probs>0.5]= 1
table(glm.pred, test$y)


fg <- pred.probs[test$y == 1]
bg <- pred.probs[test$y== 0]

# ROC Curve    
roc <- roc.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(roc)

# PR Curve
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(pr)

```


```{r  }

## sigma3
pred.probs=predict(fit.sig3,test,type="response")
glm.pred = rep(0, length(pred.probs))
glm.pred[pred.probs>0.5]= 1
table(glm.pred, test$y)


fg <- pred.probs[test$y == 1]
bg <- pred.probs[test$y== 0]

# ROC Curve    
roc <- roc.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(roc)

# PR Curve
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(pr)

```



```{r  }

## sigma4
pred.probs=predict(fit.sig3,test,type="response")
glm.pred = rep(0, length(pred.probs))
glm.pred[pred.probs>0.5]= 1
table(glm.pred, test$y)


fg <- pred.probs[test$y == 1]
bg <- pred.probs[test$y== 0]

# ROC Curve    
roc <- roc.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(roc)

# PR Curve
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
plot(pr)

```






