---
title: "model_atac"
author: "Yunqi Yang"
date: "8/11/2020"
output: html_document
---

```{r  }

load('/Users/nicholeyang/Desktop/gene_level_fine_mapping/data/train_all.RData')

```


```{r  }

head(train_all_sig)

```


## correlation by distance

Comment: ATAC features are mostly concentrated in the range < 5e5. 
```{r  }

par(mfrow = c(1,2))
plot(train_all_sig$tss_dist_to_snp, train_all_sig$correlation, xlab = 'distance', ylab = 'correlation')

plot(train_all_sig[train_all_sig$correlation>0.4, ]$tss_dist_to_snp, train_all_sig[train_all_sig$correlation>0.4, ]$correlation, xlab = 'distance', ylab = 'correlation')

```


## direct regress correlation on y  

Complete separation? 

https://courses.ms.ut.ee/MTMS.01.011/2018_spring/uploads/Main/GLM_slides_7_binary_response_ii.pdf

```{r  }

fit1 = glm(y ~ correlation, data = train_all_sig, family = "binomial")
fit2 = glm(y ~ correlation + tss_dist_to_snp, data = train_all_sig, family = "binomial")

summary(fit1)
summary(fit2)


#plot(fit1$fitted.values)
#plot(fit2$fitted.values)
```


## Try different threshold and make it binary
```{r  }

threshold = c(0.4, 0.5, 0.6, 0.7, 0.8, 0.9)
prop_pos = rep(NA, length(threshold))
prop_neg = rep(NA, length(threshold))

for (i in 1:length(threshold)){
  
  train_all_sig$atac = ifelse(train_all_sig$correlation > threshold[i], 1, 0)
  
  dat_pos = train_all_sig[train_all_sig$y == 1, ]
  dat_neg = train_all_sig[train_all_sig$y == 0, ]
  
  prop_pos[i] = sum(dat_pos$atac == 1)/dim(dat_pos)[1]
  prop_neg[i] = sum(dat_neg$atac == 1)/dim(dat_neg)[1]
  
  fit = glm(y ~ atac, data = train_all_sig, family = "binomial")
  print(summary(fit))
  
}


```


## plot the proprotion of atac features using different threshold 

Comment: shouldn't use a too stringent threshold. Maybe 0.5-0.8. 

```{r  }

prop_pos
prop_neg
prop_pos/prop_neg

prop_pos * dim(dat_pos)[1]
prop_neg * dim(dat_neg)[1]

```


```{r  }

plot(threshold, prop_pos, type = 'b', ylab = 'proportion')
points(threshold, prop_neg, type = 'b', col = 'red')
legend('topright', legend =  c('positive', 'negative'), col = c('black', 'red'), lty = 1)


```



## summary the binary feature by distance 

Comment: seems that when the distance is more close, the ATAC feature is more abundant. 
```{r  }

summary_by_dist = function(cor_threshold, dist_range){
  
  train_all_sig$atac = ifelse(train_all_sig$correlation > cor_threshold, 1, 0)
  
  prop_pos_by_dist = rep(NA, length(dist_range)-1)
  prop_neg_by_dist = rep(NA, length(dist_range)-1)
  
  
  for (i in 1:(length(dist_range)-1)){
    sub_dat = train_all_sig[train_all_sig$tss_dist_to_snp %in% seq(dist_range[i], dist_range[i+1], by = 1), ]
    sub_pos = sub_dat[sub_dat$y == 1, ]
    sub_neg = sub_dat[sub_dat$y == 0, ]
  
    prop_pos_by_dist[i] = sum(sub_pos$atac == 1)/dim(sub_pos)[1]
    prop_neg_by_dist[i] = sum(sub_neg$atac == 1)/dim(sub_neg)[1]
  }
  return(list(pos = prop_pos_by_dist, neg = prop_neg_by_dist))
}

```

```{r  }

res1 = summary_by_dist(0.5, dist_range = c(0, 1e4, 5e4, 1e5, 2e5, 3e5, 4e5, 5e5, 3e6))
res2 = summary_by_dist(0.8, dist_range = c(0, 1e4, 5e4, 1e5, 2e5, 3e5, 4e5, 5e5, 3e6))

par(mfrow = c(1,2))
plot(res1$pos, type = 'b', ylab = 'proportion', main = 'threshold=0.5')
points(res1$neg, type = 'b', col = 'red')
legend('topright', legend =  c('positive', 'negative'), col = c('black', 'red'), lty = 1, cex = 0.6)


plot(res2$pos, type = 'b', ylab = 'proportion', main = 'threshold=0.8')
points(res2$neg, type = 'b', col = 'red')
legend('topright', legend =  c('positive', 'negative'), col = c('black', 'red'), lty = 1, cex = 0.6)

```




```{r  }

res1 = summary_by_dist(0.5, dist_range = c(0, 1e4, 5e4, 1e5, 3e6))
res2 = summary_by_dist(0.8, dist_range = c(0, 1e4, 5e4, 1e5, 3e6))

par(mfrow = c(1,2))
plot(res1$pos, type = 'b', ylab = 'proportion', ylim = c(0, 0.15),  main = 'threshold=0.5')
points(res1$neg, type = 'b', col = 'red')
legend('topright', legend =  c('positive', 'negative'), col = c('black', 'red'), lty = 1, cex = 0.6)


plot(res2$pos, type = 'b', ylab = 'proportion', ylim = c(0, 0.03), main = 'threshold=0.8')
points(res2$neg, type = 'b', col = 'red')
legend('topright', legend =  c('positive', 'negative'), col = c('black', 'red'), lty = 1, cex = 0.6)

```